---
title: "First Beamer!"
author: "Kenna Rewcastle"
date: "1/30/2018"
output: 
  beamer_presentation: 
    colortheme: beaver
    incremental: yes
    theme: Madrid
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```



## Second slide

- First incremental bullet point
- Second incremental bullet point
- Third incremental bullet point 

## Third slide

>- first simultaneous bullet point
>- second simultaneous bullet point
>- third simultaneous bullet point

## Let's see some R
```{r, echo=FALSE}
library(ggplot2)
var1<-runif(10)
var2<-runif(10)
qplot(x=var1,y=var2)
```

## Colored bullet points
>- Collection Mode
>- \color{gray} Spatial and Temporal
>- \color{red} Taxonomic Bias
>- \color{gray} Sampling \color{black}
>- more stuff

## Pulling images
\begin{picture}(320,250)
\put(85,75){\includegraphics[height=6cm] {celtic.jpg}}
\put(00,60){\begin{minipage}[t]{1\linewidth} {Choose a point on the unit circle. Connect it to 
the origin with a line of length one, and denote the angle between that line 
and the horizontal coordinate axis by $\theta$.}
\end{minipage}}
\end{picture}