---
title: "Homework 4"
author: "Kenna Rewcastle"
date: "2/7/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.

```{r}
x<-1.1
a<-2.2
b<-3.3
```

a) x^a^b

```{r}
z<-x^(a^b)
print(z)
```

b) (x^a)^b

```{r}
z<-(x^a)^b
print(z)
```

c) 3x^3 + 2x^2 + 1

```{r}
z<-3*x^3 + 2*x^2 + 1
print(z)
```

d) The digit in the second place of z.

```{r}
Z<-as.integer(substring(z, seq(nchar(z)), seq(nchar(z))))
Z_use<-na.omit(Z)
print(Z_use[2])
```

2.Using the rep and seq functions, create the following vectors:

a) (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)

```{r}
a1<-seq(1:8)
a2<-seq(from=7,to=1)
a<-c(a1,a2)
print(a)
```

b) (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)

```{r}
b1<-seq(1:5)
b<-rep(b1,times=b1)
print(b)
```

c) (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

```{r}
c1<-seq(1:5)
c2<-seq(from=5,to=1)
c<-rep(c2,times=c1)
print(c)
```

3. Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates.

```{r}
myVect<-runif(n=2) # create a vector of 2 random numbers drawn from a uniform distriubtion

cart2pol <- function(vect) # fucntion that converts cartesian coordintes to polar coordinates
{
  x<-vect[1]
  y<-vect[2]
  r <- sqrt(x^2 + y^2)
  t <- atan(y/x)

  pol<-c(r,t)
  print(pol)
}

cart2pol(vect=myVect) # converts my cartesian coordinates to polar coordinates, output= polar coordinates
```

4. Suppose that queue <- c("sheep", "fox", "owl", "ant") and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as:

a) the serpent arrives

```{r}
queue<-c("sheep", "fox", "owl", "ant")
queue<-c(queue,"serpent")
print(queue)
```

b) the sheep enters the ark

```{r}

```

