---
title: "Homework 4"
author: "Kenna Rewcastle"
date: "2/7/2018"
output: html_document
---
1. Suppose x = 1.1, a = 2.2, and b = 3.3. Assign each expression to the value of the variable z and print the value stored in z.

```{r}
x<-1.1
a<-2.2
b<-3.3
```

a) x^a^b

```{r}
z<-x^(a^b)
print(z)
```

b) (x^a)^b

```{r}
z<-(x^a)^b
print(z)
```

c) 3x^3 + 2x^2 + 1

```{r}
z<-3*x^3 + 2*x^2 + 1
print(z)
```

d) The digit in the second place of z.

```{r}
Z<-as.integer(substring(z, seq(nchar(z)), seq(nchar(z))))
Z_use<-na.omit(Z)
print(Z_use[2])
```

2.Using the rep and seq functions, create the following vectors:

a) (1,2,3,4,5,6,7,8,7,6,5,4,3,2,1)

```{r}
a1<-seq(1:8)
a2<-seq(from=7,to=1)
a<-c(a1,a2)
print(a)
```

b) (1,2,2,3,3,3,4,4,4,4,5,5,5,5,5)

```{r}
b1<-seq(1:5)
b<-rep(b1,times=b1)
print(b)
```

c) (5,4,4,3,3,3,2,2,2,2,1,1,1,1,1)

```{r}
c1<-seq(1:5)
c2<-seq(from=5,to=1)
c<-rep(c2,times=c1)
print(c)
```

3. Create a vector of two random uniform numbers. In a spatial map, these can be interpreted as x and y coordinates that give the location of an individual (such as a marked forest tree in a plot that has been mapped). Using one of R’s inverse trigonometry functions (asin(), acos(), or atan()), convert these numbers into polar coordinates.

```{r}
myVect<-runif(n=2) # create a vector of 2 random numbers drawn from a uniform distriubtion

cart2pol <- function(vect) # fucntion that converts cartesian coordintes to polar coordinates
{
  x<-vect[1]
  y<-vect[2]
  r <- sqrt(x^2 + y^2)
  t <- atan(y/x)

  pol<-c(r,t)
  print(pol)
}

cart2pol(vect=myVect) # converts my cartesian coordinates to polar coordinates, output= polar coordinates
```

4. Suppose that queue <- c("sheep", "fox", "owl", "ant") and that queue represents the animals that are lined up to enter Noah’s Ark, with the sheep at the front of the line. Using R expressions, update the queue successively as:

a) the serpent arrives

```{r}
queue<-c("sheep", "fox", "owl", "ant")
queue<-c(queue,"serpent")
print(queue)
```

b) the sheep enters the ark

```{r}
queue<-queue[2:5]
print(queue)
```

c) the donkey arrives and talks his way to the front of the line

```{r}
queue<-c("donkey",queue)
print(queue)
```

d) the serpent gets impatient and leaves

```{r}
queue<-queue[queue != "serpent"]
print(queue)
```

e) the owl gets bored and leaves

```{r}
queue<-queue[queue != "owl"]
print(queue)
```

f) the aphid arrives and the ant invites him to cut in line

```{r}
queue<-c(queue[1:2],"aphid",queue[3])
print(queue)
```

g) Finally, determine the position of the aphid in the line.

```{r}
which(queue == "aphid")
```

5. Use R to create a vector of all of the integers from 1 to 100 that are not divisible by 2, 3, or 7.

```{r}
start5<-seq(1:100)
final5<-start5[!start5%%2==0 & !start5%%3==0 & !start5%%7==0] # %% gives you the remainder
print(final5)
```

6. Create a vector z of 1000 random uniform numbers.

```{r}
z<-runif(1000)
```

a) Create a vector that contains 3 numbers: the proportion of the numbers in z that are less than 0.10, greater than 0.90, and between 0.45 and 0.55.

```{r}
a<-z[z<0.10]
b<-z[z>0.90]
c<-z[z>0.45 & z<0.55]

final_vect<-c((length(a)/1000),(length(b)/1000),(length(c)/1000))
print(final_vect)
```

b) Making successive copies of z, transform your vector of uniform numbers in the following ways.
c) For each case calculate your vector of 3 numbers to get the new proportions.

b1) log (base 10) of z

```{r}
z1<-log10(z) # transform z

a1<-z1[z1<0.10]
b1<-z1[z1>0.90]
c1<-z1[z1>0.45 & z1<0.55]

final_vect1<-c((length(a1)/1000),(length(b1)/1000),(length(c1)/1000)) # calculates proportions
print(final_vect1)
```

b2) z^2

```{r}
z2<-z^2

a2<-z2[z2<0.10]
b2<-z2[z2>0.90]
c2<-z2[z2>0.45 & z2<0.55]

final_vect2<-c((length(a2)/1000),(length(b2)/1000),(length(c2)/1000))
print(final_vect2)
```

b3) e^z

```{r}
z3<-exp(z) 

a3<-z3[z3<0.10]
b3<-z3[z3>0.90]
c3<-z3[z3>0.45 & z3<0.55]

final_vect3<-c((length(a3)/1000),(length(b3)/1000),(length(c3)/1000))
print(final_vect3)
```

b4) square root of z

```{r}
z4<-sqrt(z)

a4<-z4[z4<0.10]
b4<-z4[z4>0.90]
c4<-z4[z4>0.45 & z4<0.55]

final_vect4<-c((length(a4)/1000),(length(b4)/1000),(length(c4)/1000))
print(final_vect4)
```


