---
title: "Homework 7 and 8"
author: "Kenna Rewcastle"
date: "3/21/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Creating Fake Data Sets to Explore Hypotheses

The purpose of this exercise is to statistically explore an experimental design using simulated data before implenting the design. We will simulate data that would be expected if a specific hypothesis were true. 

In this case, we assume that the presence of roots and mycorrhizae stimulates microbial assimilation of carbon substrates. The presence of these rhizosphere components has a larger effect on the assimilation of leaf substrates than on the assimilation of labile starch substrates. Tree type does not affect assimilation of these substrates.

### Simulating data for each treatment group.

The response variable in this context is the $\delta$ ^13^C signal in the phospholipid components of microbial cell membranes. Background, naturally occurring $\delta$ ^13^C values for soil microbes is approximately -25 &permil;, and values closer to 0 &permil; $\delta$ ^13^C can be expected when microbes are actively breaking down and assimilating labeled substrates to the greatest extent. For each of six treatment groups (2 substrate types x 3 rhizosphere manipulation core types) there are 12 replicates.

#### Setting up a data frame.
```{r}
ID<-seq(from=1,to=72) # creates a unique ID for each sample
core<-rep(1:3,each=12,times=2) # string of core treatment IDs
label<-rep(c("L","S"),each=36) # string of substrate type IDs

```
